Type Error Telemetry at Scale: Pseudonymized and Stochiastic

### 2023-09-20 programming23/paper.tex
resolve a few comments
type error popularity notes

%% out/negative-edit-range.txt
%% actual value is NOT negative, but large positive over 4 billion
%% example: 4294967293
Third, the edit ranges are negative in a few hundred records~(1,533).
This is likely due to the user deleting a large block of code.
Since the issue affects so few records, we omit the negative ranges
from analysis (but, we do use other parts of the records; for example
the timestamps influence~\cref{f:records-per-hour}).
%% (\FILL{} Alan, double-check reasoning)
%% Seems fine by me

@@

% That seems way off, not sure we're seeing so many files :/

@@

% I think this is too negative!
%% These adoption numbers are concerning.
%% Creators either do not know about type analysis, or they choose not
%% to use it in the code they are currently maintaining.
%% Uncovering the reason for low adaption is an important topic for future work on
%% Luau that has implications for the gradual typing community~(\cref{s:discussion}).

@@

The data does not support this explanation.
On average, the mode-upgrades in the data resulted in 3 additional
type errors (stddev 7, median 0).
The worst-case increase was 57 type errors, which is very high
but exceptional.
In another exception, upgrading modes removed three type errors
% (presumably due to the increased precision of \mstrict{} mode
% \FILL{} Alan, sounds plausible?).
% My guess as to what;s going on
(possibly due to ther edits being rolled in with the
mode change).

@@

On a coarser level, \cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
For each error type that ever overlaps, the table
shows what percentage it takes relative to all overlaps.

The table does not report \mnocheck{} because all its type errors are
by definition \code{SyntaxError}s.

@@

\begin{itemize}
  \item
    The most popular errors are syntax errors.
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

  \item
    \FILL{} relative mismatches between NS and S?

  \item
    \FILL{} which errors that NEVER showed up?

FILL N errors never showed up.
Nothing for \code{CodeTooComplex} or \code{UnificationTooComplex} is great news.
But what about the others? FILL

@@

Fortunately, the data rarely contains too-complex errors.
There are zero \code{UnificationTooComplex} errors and only a
handful of \code{CodeTooComplex} errors.
%% 26 total
It is entirely possible that the \code{CodeTooComplex} errors came from
a single user, as they appear in eleven records (out of 1.5 million)
spread across three sessions.
At any rate, the overall number is low enough to suggest that the internal
limits are not posing barriers to typechecker adoption.

\subsection{Analysis 2}

TODO work this into the above or below.

What can we learn from the telemetry? We have 1.5 million records from Feb 22
to April 14 (roughly 27K records per day).
They mention +70 million type errors.
(Some records point to a large number of errors. FILL median, max. Could be cascading,
which makes the 70million number uninterpretable.)
What takeaways are there??!

First,
users rarely switch type analysis mode. There are 3 modes available and users
are free to switch between them, or to mix modes in a codebase, but we rarely
detect these events:

\pct{99} of sessions use a single mode

\pct{0.2} of sessions use different modes in different modules. (More
precisely, the user switched to a module --- triggering a telemetry record ---
that used a different mode than the previous one.)

\pct{0.2} of records show a mode switch (equally split between upgrades and
downgrades, 341 and 320)

>>>>

%% out/negative-edit-range.txt
%% actual value is NOT negative, but large positive over 4 billion
%% example: 4294967293
Third, the edit ranges are negative in a few hundred records~(1,533).
This is likely due to the user deleting a large block of code.
Since the issue affects so few records, we omit the negative ranges
from analysis (but, we do use other parts of the records; for example
the timestamps influence~\cref{f:records-per-hour}).

@@

(\FILL{} doublecheck file sizes)
% That seems way off, not sure we're seeing so many files :/

@@

@@

The data does not support this explanation.
On average, the mode-upgrades in the data resulted in 3 additional
type errors (stddev 7, median 0).
The worst-case increase was 57 type errors, which is very high
but exceptional.
In another exception, upgrading modes removed three type errors
(possibly due to other edits being rolled in with the
mode change).

@@

\Cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
This table is based on the same data as the previous table~(\cref{t:type-error-survival}),
but focuses on the overall popularity of type errors
rather than their relation to edits.
The left column in each row names one kind of type error
that overlaps with an edit region.
The right column shows what percentage of all overlapping
errors the current error took up.

There are two subtables, for \mnonstrict{} and \mstrict{} mode.
The table does not report \mnocheck{} because all its type errors are
by definition \code{SyntaxError}s.

@@

\begin{itemize}
  \item
    As we noted above, the most popular errors are syntax errors
    (\pct{FILL} \mnonstrict{}, \pct{FILL} \mstrict{})..
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

  \item
    \code{TypeMismatch} is far more common in \mstrict{} (\pct{20.46})
    than \mnonstrict{} (\pct{2.44}) because \mstrict{} mode
    has additional type constraints.
    Several other errors have higher \mstrict{} percentages for the
    same reason, for example, \code{CannotInferBinaryOperation}.

  \item
    Two errors that appear in the \mstrict{} records
    never appear in \mnonstrict{}: \code{OccursCheckFailed} and \code{IncorrectGenericParameterCount}.
    We attribute these to more advanced uses of types in \mstrict{} mode.

  \item
    Four errors from the \mnonstrict{} records never appeared in \mstrict{} mode:
    \code{DuplicateTypeDefinition}, \code{OnlyTablesCanHaveMethods},
    \code{MissingUnionProperty}, and \code{FunctionDoesNotTakeSelf}.
    (\FILL{} Alan is this interesting, or just a result of having less strict data?)

  \item
    \FILL{} which errors that NEVER showed up?
    Nothing for \code{CodeTooComplex} or \code{UnificationTooComplex} is great news.
    But what about the others? FILL

@@

Fortunately, the data rarely contains too-complex errors.
None of these errors appear in edit regions~(\cref{t:type-error-count}).
For the specific case of \code{CodeTooComplex}, our telemetry tracked project-wide
counts and found only 26 of these errors, which were spread across eleven
records in three sessions.
It is entirely possible that all these errors came from one creator who
deliberately pushed the limits of type analysis.
Whatever the reason, the overall number is low enough to suggest that the
internal limits are not posing barriers to typechecker adoption.

\subsection{NEXT}
